---
title: "Does Physical Activity Differ by Marital Status?"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      library(tidyverse),
                      library(knitr),
                      library(aplore3))
```

## Overview

This project investigates whether physical activity patterns differ across marital status groups using data from the NHANES dataset. The primary goal is to determine if individuals' levels of physical activity — such as vigorous work, moderate exercise, or transportation by walking or biking — vary depending on their marital status.

## Research Question

**Is there a statistically significant difference in physical activity levels across different marital statuses?**

## Variables of Interest

-   **Marital Status**\
    Coded into six categories:
    -   Married\
    -   Widowed\
    -   Divorced\
    -   Separated\
    -   Never Married\
    -   Living Together
-   **Physical Activity Types**\
    Binary (Yes/No) indicators of whether the respondent engaged in:
    -   Vigorous Work Activity\
    -   Moderate Work Activity\
    -   Walking or Biking\
    -   Vigorous Recreational Activity\
    -   Moderate Recreational Activity

## Methodological Approach

The methodology uses **contingency tables** to examine the relationship between marital status and each type of physical activity. For each table:

-   We observe the **raw counts** of individuals participating (or not) in a given activity across marital groups.
-   We compute **row-wise proportions** to understand how physical activity behaviors differ within each marital group.
-   We apply a **Chi-squared test of independence** to assess whether the differences are statistically significant.

```{r}
# Clean and prepare the dataset
df <- nhanes %>%
  drop_na(marstat, vigwrk, modwrk, wlkbik, vigrecexr, modrecexr, sedmin) # Drops all the missing values

str(df)

```

```{r}
# Template function
analyze_by_marstat <- function(activity_var, label) {
  tab <- table(Activity = df[[activity_var]], MaritalStatus = df$marstat)
  row_prop <- round(addmargins(prop.table(tab, margin = 1), margin = 2), 3)
  
  print(kable(addmargins(tab, margin = 2), caption = paste(label, "vs Marital Status")))
  
  cat("\n Row Proportions:")
  print(kable(row_prop))

  print(chisq.test(tab))
  
}

# Run for all physical activity types
analyze_by_marstat("vigwrk", "Vigorous Work Activity")
analyze_by_marstat("modwrk", "Moderate Work Activity")
analyze_by_marstat("wlkbik", "Walking or Biking")
analyze_by_marstat("vigrecexr", "Vigorous Recreational Activity")
analyze_by_marstat("modrecexr", "Moderate Recreational Activity")

```

### Interpretation Guide

For each type of physical activity, a significant Chi-squared result (p \< 0.05) suggests a relationship between that activity and marital status - there's statistical evidence that physical activity is not independent of marital status.

Row proportions help us see which groups are more or less active, providing insight into behavioral differences that may be shaped by lifestyle, responsibilities, or social structures tied to marital status.
